Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Charles Yang, and audiences at NELS 43 for helpful comments.}

### 2013-01-17 1.tex
Pre-"revisions", with lots of material in the TODO

This dissertation has two distinct but complementary aims. 
The first is to outline the empirical scope of \emph{phonotactic theory}, the theory of speakers' knowledge of possible and impossible sounds, sound sequences, and words.
The second is to show that the core facts of this domain are compatible with a traditional view of phonotactic knowledge as independent of the lexicon, categorical, and closely related to phonological processes, but inconsistent with an increasingly popular view of phonotactics and phonotactic learning as a type of probabilistic inference over the lexicon, and therefore gradient and independent of phonological processes.\footnote{
    Throughout, the term \emph{lexicon} is used in a specific sense of the set underlying representations in some language; this is not meant to imply a position on the possibility that larger, composite linguistic representations are also stored in lexical memory.
    See \citealt{LignosInPressa} for a review of experimental evidence on this question.}

@@

\noindent
\citeauthor{PE} go on to label constraints on underlying representations as ``extensional'', and thus irrelevant to generative grammar, in contrast to ``intensional'' statements like phonological rules.
Regarding the facts about Russian obstruents discussed so far, it seems to be clear that the alternation facts are in some sense more privileged.
Were it the case that Russian speakers had not internalized a process of obstruent voicing assimilation, the only alternative explanation for the forms in (\ref{rovs}) is a massive system of phonologically conditioned suppletive allomorphy.
In contrast, were there no contraint on underlying obstruent voicing, however, it is not immediately obvious that \emph{anything} would be different. 
Presumably, this is what \citeauthor{PE} mean when they refer to constraints on underlying representations as ``computationally inert''.
There are also many cases where restrictions on underlying representations are specific to underlying representations and do not apply to surface representations.
For instance, hiatus is exceptionally rare in native Turkish roots; of the handful of examples in the Turkish Electronic Living Lexicon \citep{TELL}, many appear to be compounds (e.g., \emph{ısıalan} `endothermic', cf.~\emph{ısı} `heat', \emph{alan} `taker').
Yet numerous phonological processes give rise to hiatus in derived environments \citep[e.g.,][]{Kabak2007b}.
\citet{Silverman2000} observes that operations like reduplication and truncation also tend to introduce violations of constraints on underlying representations.
However, there are many facts which strongly suggest that speakers internalize phonotactic constraints.

@@

Speakers have difficulty producing \citep[e.g.,][]{Davidson2006a,Davidson2010,Rose2007,Vitevitch1997} and perceiving \citep{Dupoux1999,Kabak2007a,Massaro1983} certain types of nonce words judged to be phonotactically illicit.
For instance, \citet{GallagherInPress} finds that speakers of Quechua have difficulty repeating nonce words with multiple ejectives (e.g., [k'ap'i]), which do not occur in the language; this suggests that 
multiple ejective words are not merely absent, but actually difficult for these speakers to perceive, represent, or produce.

In English, sequences of adjacent obstruents which do not also agree in voice (e.g., \emph{a}[b.s]\emph{inth}) are quite rare within a word, and therefore a hetero-voiced obstruent cluster is a clue to the presence of a word boundary in running speech.
Infants \citep[e.g.,][]{Mattys2001b} and adults \citep{McQueen1998b,Norris1997} are both thought to use this heuristic for word recognition in experimental settings.\footnote{
    There is reason to distinguish the computations involved in using this word segmentation heuristic from those implicated by other phonotactic behaviors.
    Whatever the locus of [bnɪk]'s ill-formedness, for example, no segmentation into multiple words or morphs renders it a well-formed sequence.
    See \S\ref{ss:bh} for further discussion.}

@@

Loanword adaptation may provide further evidence for the psychological reality of phonotactic knowledge.
In Desano \citep{Kaye1974}, for instance, all underlying representations (URs) are either totally oral (e.g., [yaha] `to hear') or totally nasal (e.g., [ñãhã] `to enter'), and loanwords are made to conform to this generalization: Portuguese \emph{martelo} `hammer' is adapted as [barateru] and Spanish \emph{naranja} `orange' as [nãnãnã].
Presumably, some component of the synchronic grammar is responsible for the fact that the restriction over native vocabulary is extended to loanwords.

%FIXME

@@

There are many other cases, however, where it is claimed that phonotactic restrictions are not extended to loanwords (e.g., \citealt{Clements1982}, \citealt[75]{Davidson1997}, \citealt{Fries1949}, \citealt{Holden1976}, \citealt{Ito1995a,Ito1995b}, \citealt[95]{Shibatani1973}, \citealt{Ussishkin2003}, \citealt{Vogt1954}; additional examples can be found throughout this dissertation).
For instance, native words in San Mateo Huave all end in a consonant, but final unstressed syllables in Spanish loanwords are never repaired by epenthesis (e.g., \emph{verde} `green' $>$ [beɾde], *[beɾdej]).
%FIXME example here
Given the limited understanding of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be a useful diagnostic.

@@

Whereas \citeauthor{Shibatani1973} argues that morpheme structure constraints are insufficient to account for speakers' phonotactic knowledge, other authors observe the tendency for structural descriptions of phonological processes to reappear among the morpheme structure constraints on the same language (e.g., \citealt{Hale1965}:297, \citealt{Kisseberth1970b}, \citealt{Postal1968}:212f.).
Russian obstruent voice assimilation, discussed above, provides an example of this type: there are no hetero-voiced obstruent clusters in either underlying or surface representations.
These two facts are tantalizingly similar, but are treated as separate if a distinction between morpheme structure constraints and phonological processes is drawn.
This is a special case of what \citet{Kisseberth1970b} calls \emph{conspiracies}, which \citet[136]{KK77} term the \emph{duplication problem}.
\citet[205f.]{Dell1973} and \citet[28f.]{Stampe1973} argue that the distinction between constraints on URs and alternations is artificial, and that these different levels of description are related by a principle now known as \emph{Stampean occultation} \citep[54]{OT}.
In a language like Russian, in which surface obstruent clusters exceptionlessly agree in voicing, there is simply no reason for the language acquisition device to posit underlying hetero-voiced obstruent clusters: obstruent voice assimilation ``occults'' underlying */kb/, for instance. Were such an underlying form posited, it would surface as [gb] in all contexts.\footnote{
    \emph{Lexicon Optimization} \citep[209]{OT} implements a form of Stampean occultation notable in that it projects all non-alternating surface segments directly into URs.
    For instance, in English, Lexicon Optimization demands underlying /ŋ/ in words like \emph{bank}, though [ŋ] could be otherwise be analyzed as an allophone of /n/ before velar consonants (e.g., \citealt[65f.]{Borowsky1986}, \citealt[85]{SPE}, \citealt[62]{Halle1985a}), eliminating /ŋ/ from the phoneme inventory.
    However, this is not core to \citeauthor{Dell1973} and \citeauthor{Stampe1973}'s insight about the relationship between surface and underlying sequence structure restrictions.
    For instance, the hypothetical \{bæŋk\} posited by Lexicon Optimization could be revised to /bænk/, and take a \emph{free ride} (in the sense of \citealt{Zwicky1970}) on the process of nasal place assimilation found elsewhere in English (see \S\ref{npa}).
    Indeed, this seems desirable, since Lexicon Optimization forces a duplication between underlying and surface constraints.
%this point is made by \citet{Chomsky1957} and \citet[69f.]{Harris1960} with respect to the principle of biuniqueness, which enforces a similar restriction.
    For instance, [ŋ] does not appear word-initially and English speakers have considerable difficulty producing it in this position \citep{Rusaw2009}.
    The allophonic analysis of [ŋ] predicts this fact, since there is no way to derive the [ŋ] allophone in onset position.
    Future work will consider the how purely allophonic relationships might be acquired.}
    %Assuming Lexicon Optimization, the only alternative is to stipulate an ad hoc constraint against onset [ŋ] \citep[39]{Jusczyk2002}.}
In chapter \ref{gaps}, it is argued that constraints described in terms of non-contrastive prosodic structures can also be derived by Stampean occultation.
In an architecture like Lexical Phonology, it is even possible to apply a process to individual underlying representations (i.e., at the ``lexical level'') to enforce constraints which are not surface-true.
Consequently, morpheme structure constraints are otiose, their empirical converage entirely subsumed by phonological processes.

@@

Under the assumptions of Stampean occultation, there is no way to express substantive constraints on underlying representations except via phonological process.
In lexical phonology, such constraints are the consequences of phonological processes operating at the ``lexical'' level, and in an Optimality Theory grammar, these are high ranked markedness constraints.\footnote{In Optimality Theory, it is often assumed that all markedness constraints are ranked above all faithfulness constraints in the ``initial state'' of the learner \citep[e.g.,][]{Smolensky1996a}.
    This admits the possibility that markedness constraints not implicated by alternations will remain undominated \citep[e.g.,][]{Coetzee2008b}.
    It is difficult to evaluate this proposal in the absence of any substantive proposal for the contents of \textsc{Con}, the universal constraint sets, but this has the potential to blur the traditional distinction between accidental and structural phonotactic gaps: any exceptionless gap which corresponds to a markedness constraint \emph{in any language} will be accorded structural, inviolable status, a quite powerful prediction.
    If \textsc{Con} is sufficiently elaborate to incorporate constraints like the *[spVp] proposed by \citet{Coetzee2008b}, it might also incorporate a constraint like *[s.w], perhaps as a subcomponent of the so-called \emph{syllable contact law} \citep[e.g.,][]{Gouskova2004,Murray1983}, which disfavors syllable contact clusters with increasing sonority 
    The *[s.w] constraint is without exception in English, yet nonce words like \emph{teeswa} [tis.wa] seem quite unobjectionable.
    Numerous other examples of this type could be adduced.
    Consequently, this specific model is rejected here.}
A corrolary is that if some phonotactic constraint is not (or cannot be) identified with a phonological process, it does not exist in the minds of speakers at all.

@@

\citet{Hayes2004b} argues that phonotactic learning occurs before lexical or phonological acquisition, and it is therefore must be independent of other types of grammatical knowledge.
This is not fully consistent with the available experimental evidence, however.

@@

Crucially, all of this occurs \emph{before} there is any evidence for phonotactic knowledge, for which the earliest evidence begins at approximately 9 months of age \citep{Friederici1993,Jusczyk1993b,Jusczyk1994}.
While there are many gaps in current understanding (and keeping in mind the usual caveats about a strong interpretation of negative results), there is no reason to think that phonotactic knowledge is acquired before non-trivial amounts of lexical and phonological acquisition.
If anything, infants's phonotactic knowledge appears to come online at approximately the same point as highly specific phonological representations including subsyllabic units and the ability to extract allophonic generalizations, as predicted by the firm link between phonological and phonotactic knowledge argued for here.

@@

>>>>

This dissertation has two distinct but complementary aims. 
The first is to outline the empirical scope of \emph{phonotactic theory}, the theory of speakers' knowledge of possible and impossible sounds, sound sequences, and words.
The second is to show that the core facts of this domain are compatible with a traditional view of phonotactic knowledge as independent of the lexicon, categorical, and closely related to phonological processes, but inconsistent with an increasingly popular view of phonotactics and phonotactic learning as a type of probabilistic inference over the lexicon, and therefore gradient and independent of phonological processes.\footnote{
    Throughout, the term \emph{lexicon} is used in a specific sense of the set underlying representations in some language; this is not meant to imply a position on the possibility that larger, composite linguistic representations are also stored in lexical memory.
    For a review of recent experimental evidence on this question, see \citealt{LignosInPressa}.}

@@

\noindent
\citeauthor{PE} go on to label constraints on underlying representations as ``extensional'', and thus irrelevant to generative grammar, in contrast to ``intensional'' statements like phonological rules.
Regarding the facts about Russian obstruents discussed so far, it seems to be clear that the alternation facts are in some sense more privileged.
Were it the case that Russian speakers had not internalized a process of obstruent voicing assimilation, the only alternative explanation for the forms in (\ref{rovs}) is a massive system of phonologically conditioned suppletive allomorphy.
In contrast, were there no constraint on underlying obstruent voicing, however, it is not immediately obvious that \emph{anything} would be different. 
Presumably, this is what \citeauthor{PE} mean when they refer to constraints on underlying representations as ``computationally inert''.
There are also many cases where restrictions on underlying representations are specific to underlying representations and do not apply to surface representations.
For instance, hiatus is exceptionally rare in native Turkish roots; of the handful of examples in the Turkish Electronic Living Lexicon \citep{TELL}, many appear to be compounds (e.g., \emph{ısıalan} `endothermic', cf.~\emph{ısı} `heat', \emph{alan} `taker').
Yet numerous phonological processes give rise to hiatus in derived environments \citep[e.g.,][]{Kabak2007b}.
\citet{Silverman2000} observes that operations like reduplication and truncation also tend to introduce violations of constraints on underlying representations.
However, there are many facts which strongly suggest that speakers internalize phonotactic constraints.

@@

Speakers have difficulty producing \citep[e.g.,][]{Davidson2006a,Davidson2010,Rose2007,Vitevitch1997} and perceiving \citep{Dupoux1999,Kabak2007a,Massaro1983} certain types of nonce words judged to be phonotactically illicit.
For instance, \citet{GallagherInPress} finds that speakers of Quechua have difficulty repeating nonce words with multiple ejectives (e.g., [k'ap'i]), which do not occur in the language;
in other words, multiple ejective sequences are not merely absent, but also difficult for these speakers to repeat.

In English, sequences of adjacent obstruents which do not also agree in voice (e.g., \emph{a}[b.s]\emph{inth}) are quite rare within a word, and therefore a hetero-voiced obstruent cluster is a clue to the presence of a word boundary in running speech.
Infants \citep[e.g.,][]{Mattys2001b} and adults \citep{McQueen1998b,Norris1997} are thought to use this heuristic for word recognition in experimental settings.\footnote{
    There is reason to distinguish the computations involved in using this word segmentation heuristic from those implicated by other phonotactic behaviors.
    Whatever the locus of [bnɪk]'s ill-formedness, for example, no segmentation into multiple words or morphs renders it a well-formed sequence.
    See \S\ref{ss:bh} for further discussion.}

@@

Loanword adaptation may provide further evidence for the grammatical relevance of phonotactic knowledge.
In Desano \citep{Kaye1974}, for instance, all underlying representations (URs) are either totally oral (e.g., [yaha] `to hear') or totally nasal (e.g., [ñãhã] `to enter'), and loanwords are made to conform to this generalization: Portuguese \emph{martelo} `hammer' is adapted as [barateru] and Spanish \emph{naranja} `orange' as [nãnãnã].
It is not a stretch to imagine that some component of the synchronic grammar is responsible for the fact that the restriction over native vocabulary is extended to loanwords.
% FIXME discuss evidence that Desano has nasal spreading
% FIXME mention \citet{Jacobs2000}

@@

There are many other cases, however, where it is claimed that phonotactic restrictions are not extended to loanwords (e.g., \citealt{Clements1982}, \citealt[75]{Davidson1997}, \citealt{Fries1949}, \citealt{Holden1976}, \citealt{Ito1995a,Ito1995b}, \citealt[95]{Shibatani1973}, \citealt{Ussishkin2003}, \citealt{Vogt1954}; additional examples can be found throughout this dissertation).
For instance, native words in San Mateo Huave all end in a consonant, but final unstressed syllables in Spanish loanwords are never repaired by epenthesis (e.g., \emph{verde} `green' $>$ [beɾde], *[beɾdej]).
%FIXME example here
Given the limited understanding of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be a useful diagnostic.
% FIXME mention hyman
% peperkamp & dupoux 2003
% peperkamp 2005

@@

Whereas \citeauthor{Shibatani1973} argues that morpheme structure constraints are insufficient to account for speakers' phonotactic knowledge, other authors observe the tendency for structural descriptions of phonological processes to reappear among the morpheme structure constraints on the same language (e.g., \emph{SPE}:382, \citealt{Hale1965}:297, \citealt{Kisseberth1970b}, 
%\citealt{PasterInPress},
\citealt{Postal1968}:212f., \citet{Stanley1967}:401).
Russian obstruent voice assimilation, discussed above, provides an example of this type: there are no hetero-voiced obstruent clusters in either underlying or surface representations.
These two facts are tantalizingly similar, but are treated as separate if a distinction between morpheme structure constraints and phonological processes is drawn.
This is a special case of what \citet{Kisseberth1970b} calls \emph{conspiracies}, which \citet[136]{KK77} term the \emph{duplication problem}.
\citet[205f.]{Dell1973} and \citet[28f.]{Stampe1973} argue that the distinction between constraints on URs and alternations is artificial, and that these different levels of description are related by a principle now known as \emph{Stampean occultation} \citep[54]{OT}.
In a language like Russian, in which surface obstruent clusters exceptionlessly agree in voicing, there is simply no reason for the language acquisition device to posit underlying hetero-voiced obstruent clusters: obstruent voice assimilation ``occults'' underlying */kb/, for instance. Were such an underlying form posited, it would surface as [gb] in all contexts.\footnote{
    \emph{Lexicon Optimization} \citep[209]{OT} implements a form of Stampean occultation notable in that it projects all non-alternating surface segments directly into URs.
    For instance, in English, Lexicon Optimization demands underlying /ŋ/ in words like \emph{bank}, though [ŋ] could be otherwise be analyzed as an allophone of /n/ before velar consonants (e.g., \citealt[65f.]{Borowsky1986}, \citealt[85]{SPE}, \citealt[62]{Halle1985a}), eliminating /ŋ/ from the phoneme inventory.
    However, this is not core to \citeauthor{Dell1973} and \citeauthor{Stampe1973}'s insight about the relationship between surface and underlying sequence structure restrictions.
    For instance, the hypothetical \{bæŋk\} posited by Lexicon Optimization could be revised to /bænk/, and take a \emph{free ride} (in the sense of \citealt{Zwicky1970}) on the process of nasal place assimilation found elsewhere in English (see \S\ref{npa}).
    Indeed, this seems desirable, since Lexicon Optimization forces a duplication between underlying and surface constraints.
%this point is made by \citet{Chomsky1957} and \citet[69f.]{Harris1960} with respect to the principle of biuniqueness, which enforces a similar restriction.
    For instance, [ŋ] does not appear word-initially and English speakers have considerable difficulty producing it in this position \citep{Rusaw2009}.
    The allophonic analysis of [ŋ] predicts this fact, since there is no way to derive the [ŋ] allophone in onset position.
    Future work will consider the how purely allophonic relationships might be acquired.}
    %Assuming Lexicon Optimization, the only alternative is to stipulate an ad hoc constraint against onset [ŋ] \citep[39]{Jusczyk2002}.}
In chapter \ref{gaps}, it is argued that constraints described in terms of non-contrastive prosodic structures can also be derived by Stampean occultation.
In an architecture like Lexical Phonology, it is even possible to apply a process to individual underlying representations (i.e., at the ``lexical level'') to enforce constraints which are not surface-true.
Consequently, morpheme structure constraints are otiose, their empirical converage entirely subsumed by phonological processes.

@@

Under the assumptions of Stampean occultation, there is no way to express substantive constraints on underlying representations except via phonological process.
In lexical phonology, such constraints are the consequences of phonological processes operating at the ``lexical'' level, and in an Optimality Theory grammar, these are high ranked markedness constraints.\footnote{In Optimality Theory, it is often assumed that all markedness constraints are ranked above all faithfulness constraints in the ``initial state'' of the learner \citep[e.g.,][]{Smolensky1996a}.
    This admits the possibility that markedness constraints not implicated by alternations will remain undominated \citep[e.g.,][]{Coetzee2008b}.
    It is difficult to evaluate this proposal in the absence of any complete proposal for the contents of \textsc{Con}, the universal constraint sets, but this has the potential to blur the traditional distinction between accidental and structural phonotactic gaps: any exceptionless gap which corresponds to a markedness constraint \emph{in any language} will be accorded structural, inviolable status, a quite powerful prediction.
    If \textsc{Con} is sufficiently elaborate to incorporate constraints like the *[spVp] proposed by \citet{Coetzee2008b}, it might also incorporate a constraint like *[s.w], perhaps as a subcomponent of the so-called \emph{syllable contact law} \citep[e.g.,][]{Gouskova2004,Murray1983}, which disfavors syllable contact clusters with increasing sonority 
    The *[s.w] constraint is without exception in English, yet nonce words like \emph{teeswa} [tes.wa] seem quite unobjectionable.
    % FIXME tilwa?
    Numerous other examples of this type could be adduced.
    Consequently, this specific model is rejected here.}
A corrolary is that if some phonotactic constraint is not (or cannot be) identified with a phonological process, it does not exist in the minds of speakers at all.

@@

\citet{Hayes2004b} argues that phonotactic learning occurs before lexical or phonological acquisition, and it therefore must be independent of other types of grammatical knowledge.
This is not fully consistent with the available experimental evidence, however.

@@

Crucially, all of this occurs \emph{before} there is any evidence for phonotactic knowledge, for which the earliest evidence begins at approximately 9 months of age \citep{Friederici1993,Jusczyk1993b,Jusczyk1994}.
% Sketch what this is.
While there are many gaps in current understanding (and keeping in mind the usual caveats about a strong interpretation of negative results), there is no reason to think that phonotactic knowledge is acquired before non-trivial amounts of lexical and phonological acquisition.
If anything, infants's phonotactic knowledge appears to come online at approximately the same point as highly specific phonological representations including subsyllabic units and the ability to extract allophonic generalizations, as predicted by the firm link between phonological and phonotactic knowledge argued for here.

@@

Chapter \ref{conclusions} summarizes the findings and describes future directions.

### 2013-01-17 2.tex
Pre-"revisions", with lots of material in the TODO

One of the major  for the development of stochastic models o
motivations for the 
Recent developments in the theory of phonotactic knowledge are motivated by the claim that speakers' wordlikeness intuitions are inherently \emph{gradient}, i.e., that such intuitions are more granular than a mere binary contrast between ``possible'' and ``impossible'' words.

@@

Each step of this procedure is dubious, however.
First, speakers have difficulty perceiving \citep{Berent2007a,Brown1956,Dupoux1999,Kabak2007a} and producing \citep{Davidson2005,Davidson2006a,Davidson2006b,Davidson2010,GallagherInPress,Rose2007,Vitevitch1998,Vitevitch2005} phonotactically illicit non-words, suggesting that speakers' ability to faithfully parse illicit representations is at best quite limited.
Secondly, the computation of a scalar value serves no further purpose than to provide for gradient well-formedness judgements; even if this is the result of probabilistic parsing \citep[e.g.,][]{Coleman1997}, 
it need not be computed overtly.\footnote{
    An objection might be made here on evolutionary grounds. 
    It is still quite mysterious why the human language endowment includes constraints on pronominal binding, for instance, but these constraints are implicated in everyday language use.
    Far more bizarre is the suggestion that the linguistic endowment includes mechanisms only used in certain experimental tasks.}
Next, speakers must be able to consciously access and report the magnitude of this value (it must be \emph{cognitively penetrable} in the sense of \citealt{Pylyshyn1984}), an ability which is limited in many other domains.
Finally, \citet{Sprouse2011} argues that speakers do not (or cannot) scale well-formedness values to satisfy the assumptions of magnitude estimation, a type of gradient rating task.
%\citep{Bard1996}

It is informative to compare this baroque architecture to a model of what is necessary to make a binary well-formedness judgement. When presented with a linguistic item in a judgement task, the grammar attempts to assign a parse. Speakers then access whether or not parsing was successful. 
There are reasons to think that parsing of ungrammatical structures does in fact result in a ``crash'': whereas syntactic priming increases the acceptability of grammatical structures \citep{Luka2005}, ungrammatical structures show no priming effects \citep{Sprouse2007b}. 
As priming of linguistic structures is thought to implicate shared representations in memory, ungrammatical structures may not be stored in linguistic memory. 
%This cannot be reduced to the fact that ungrammatical structures fail to denote, since well-formed but non-existent structures do produce facilitory priming \citep[e.g.,][]{Longtin2005}. 
The fact that requests for repetition and clarification are ubiquituous in spontaneous speech illustrates further that speakers are frequently aware when parsing has failed.
Consequently, a great deal of evidence is needed to reject this simple model in favor of the gradient grammaticality architecture.

@@

It is not that categorical models have been ignored by the literature on wordlikeness modeling, but rather that they have not been compared. 
\citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated with a probabilistic measure of the well-formedness of the constituent syllables. 
Unfortunately, no attempt is made to control for the well-formedness of syllable contact clusters in these words: some of the stimuli have medial consonant clusters containing both voiced and voiceless obstruents (e.g., [ɡaɪbsaɪk]), something which is exceptionally rare in English simplex words.
Similarly, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of phonotactic constraints proposed by \citet{Clements1983}, first transform these constraints, many of which are exceptionless, into probabilities. 
While this is consistent with their claim, that ``the ability to model gradient intuitions [are] an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, this principle would preclude any attempt to test the hypotheses that underlies it.

@@

>>>>

One of the major  for the development of stochastic models o
%FIXME
motivations for the 
Recent developments in the theory of phonotactic knowledge are motivated by the claim that speakers' wordlikeness intuitions are inherently \emph{gradient}, i.e., that such intuitions are more granular than a mere binary contrast between ``possible'' and ``impossible'' words.

@@

Each step of this procedure is dubious, however.
First, speakers have difficulty perceiving \citep{Berent2007a,Brown1956,Dupoux1999,Kabak2007a} and producing \citep{Davidson2005,Davidson2006a,Davidson2006b,Davidson2010,GallagherInPress,Rose2007,Vitevitch1998,Vitevitch2005} phonotactically illicit non-words, suggesting that speakers' ability to faithfully parse illicit representations is at best quite limited.
Secondly, the computation of a scalar value serves no further purpose than to provide for gradient well-formedness judgements; even if this is the result of probabilistic parsing \citep[e.g.,][]{Coleman1997}, 
it need not be computed overtly.\footnote{
    An objection might be made here on evolutionary grounds. 
    It is still quite mysterious why the human language endowment includes constraints on pronominal binding, for instance, but these constraints are implicated in everyday language use.
    Far more bizarre is the suggestion that the linguistic endowment includes mechanisms only used in certain experimental tasks.}
Next, speakers must be able to consciously access and report the magnitude of this value (it must be \emph{cognitively penetrable} in the sense of \citealt{Pylyshyn1984}), an ability which is limited in many other domains.
Finally, \citet{Sprouse2011} argues that speakers do not (or cannot) scale well-formedness values to satisfy the assumptions of magnitude estimation, a type of gradient rating task.
%\citep{Bard1996}
% FIXME short discussion of the utility of gradient judgements in syntax

It is informative to compare this baroque architecture to a model of what is necessary to make a binary well-formedness judgement. When presented with a linguistic item in a judgement task, the grammar attempts to assign a parse. Speakers then access whether or not parsing was successful. 
There are reasons to think that parsing of ungrammatical structures does in fact result in a ``crash'': whereas syntactic priming increases the acceptability of grammatical structures \citep{Luka2005}, ungrammatical structures show no priming effects \citep{Sprouse2007b}. 
As priming of linguistic structures is thought to implicate shared representations in memory, ungrammatical structures cannot be stored in linguistic memory. 
%This cannot be reduced to the fact that ungrammatical structures fail to denote, since well-formed but non-existent structures do produce facilitory priming \citep[e.g.,][]{Longtin2005}. 
The fact that requests for repetition and clarification are ubiquituous in spontaneous speech illustrates further that speakers are frequently aware when parsing has failed.
Consequently, a great deal of evidence is needed to reject this simple model in favor of the gradient grammaticality architecture.

@@

It is not that categorical models have been ignored by the literature on wordlikeness modeling, but rather that they have not been compared. 
\citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated with a probabilistic measure of the well-formedness of the constituent syllables. 
Unfortunately, no attempt is made to control for the well-formedness of syllable contact clusters in these words: some of the stimuli have medial consonant clusters containing both voiced and voiceless obstruents (e.g., [ɡaɪbsaɪk]), something which is exceptionally rare in English simplex words.
Similarly, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of phonotactic constraints proposed by \citet{Clements1983}, first transform these constraints, many of which are exceptionless, into probabilities. 
While this is consistent with their claim, that ``the ability to model gradient intuitions [is] an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, this principle would preclude any attempt to test the hypotheses that underlies it.

% FIXME make this work: As \citet{Newmeyer2007} writes, ``the idea that categoricity is not represented in the data itself is a truism. Whether distinctions of grammaticality (as opposed to acceptability) are binary is a difficult question.'' (398)

@@

% FIXME note that it might still be useful to use continuous judgements to gather things, since it might have more statistical power (sprouse)
% featherston 2007
